Causal Bandits: Learning Good Interventions via Causal Inference

### 2016-05-14 causal-bandit/nips_2016/discussion.tex
first draft

\vspace{-0.4cm} 
\paragraph{Causal Models with Non-Observable Variables}
If we assume knowledge of the conditional \textit{interventional} distributions $P_a$ our analysis applies unchanged to the case of causal models with 
non-observable variables. For example, if we had access to a data set of \textit{experiments} in which the reward variable $Y$ was not 
available from which to build estimates of $P_a$. Some of the interventional distributions $P_a$ may be non-identifiable meaning we can not obtain prior estimates for $P_a$ from 
even an infinite amount of observational data. Even if all variables are observable and the graph is known, if the conditional distributions are unknown, then Algorithm
\ref{alg:general} cannot be used. Estimating these quantities while simultaneously minimising the simple regret is an interesting and challenging open problem.

% In this case, some conditional distributions may be non-identifiable. 
% The corresponding actions can be immediately added to the set $A$ prior to collecting any data. 
% We can then use the same algorithm as in the case where there are no latent variables, except that we will have to use the more general do-calculus rather than simply adjusting for the parents to write the expression for each action in terms of observational data.
% Combining our estimation techniques with insights from \citet{Bareinboim2015} for handling unobserved confounders would be worth investigation.

>>>>

\vspace{-0.4cm} 
\paragraph{Causal Models with Non-Observable Variables}
If we assume knowledge of the conditional \textit{interventional} distributions $\Pn{a}{\parents{Y}}$ our analysis applies unchanged to the case of causal models with 
non-observable variables. Some of the interventional distributions may be non-identifiable meaning we can not obtain prior estimates for $\Pn{a}{\parents{Y}}$ from 
even an infinite amount of observational data. Even if all variables are observable and the graph is known, if the conditional distributions are unknown, then Algorithm
\ref{alg:general} cannot be used. Estimating these quantities while simultaneously minimising the simple regret is an interesting and challenging open problem.

% For example, if we had access to a data set of \textit{experiments} in which the reward variable $Y$ was not available from which to build estimates of $P_a$.
% In this case, some conditional distributions may be non-identifiable. 
% The corresponding actions can be immediately added to the set $A$ prior to collecting any data. 
% We can then use the same algorithm as in the case where there are no latent variables, except that we will have to use the more general do-calculus rather than simply adjusting for the parents to write the expression for each action in terms of observational data.
% Combining our estimation techniques with insights from \citet{Bareinboim2015} for handling unobserved confounders would be worth investigation.

### 2016-05-14 causal-bandit/nips_2016/experiments.tex
condences figures, minor changes to section 4
first draft

We compare Algorithms \ref{alg:simple} and \ref{alg:general} with Successive Elimination on the parallel bandit problem
under a variety of conditions, including where the importance weighted estimator used by Algorithm \ref{alg:general} is not truncated,
which is justified in this setting by Remark \ref{rem:truncate}. Throughout we use a model in which $Y$ depends only on a single variable (this is unknown to the algorithms). 
\eq{
Y_t \sim \begin{cases}
\bernoulli(\frac{1}{2}+\epsilon) & \text{if } X_1 = 1 \\
\bernoulli(\frac{1}{2}-\frac{q_1}{1-q_1}\epsilon) & \text{otherwise}\,.
\end{cases}
}

This leads to an expected reward of $\frac{1}{2}+\epsilon$ for $do(X_1=1)$, $\left(\frac{1}{2}-\frac{q_1}{1-q_1}\epsilon\right)$ for $do(X_1=0)$ and $\frac{1}{2}$ for all other actions. We set $\boldsymbol{q} = 0$ for $q_i \leq m$ and $\frac{1}{2}$ otherwise. Note that changing $m$ and thus $\boldsymbol{q}$ has no effect on the reward distribution.

@@

\begin{figure}[h]
\centering
\includegraphics[width=.5\textwidth]{exp_regret_vs_m_N50_T400_s10000_20160205_2230}
\caption{Simple regret vs $m(\boldsymbol{q})$ for fixed horizon $T=400$ and number of variables $N = 50$}
\label{fig:simple_vs_m}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=.5\textwidth]{exp_regret_vs_T_N50_a4_s10000_20160204_1831}
\caption{Simple regret vs horizon, $T$, with $N = 50$, $m=2$ and $\epsilon = \sqrt{\frac{N}{8T}}$}
\label{fig:simple_vs_T_vary_epsilon}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=.5\textwidth]{exp_regret_vs_T_N50_s10000_20160206_0533}
\caption{Simple regret vs horizon, $T$, with $N = 50$, $m=2$ and fixed $\epsilon = .3$}
\label{fig:simple_vs_T}
\end{figure}

>>>>

We compare Algorithms \ref{alg:simple} and \ref{alg:general} with Successive Elimination on the parallel bandit problem
under a variety of conditions, including where the importance weighted estimator used by Algorithm \ref{alg:general} is not truncated,
which is justified in this setting by Remark \ref{rem:truncate}. Throughout we use a model in which $Y$ depends only on a single variable $X_1$ (this is unknown to the algorithms). $Y_t \sim \bernoulli(\frac{1}{2}+\epsilon)$ if $X_1=1$ and $Y_t \sim \bernoulli(\frac{1}{2}-\epsilon')$ otherwise, where $\epsilon' = q_1\epsilon/(1-q_1)$. This leads to an expected reward of $\frac{1}{2}+\epsilon$ for $do(X_1=1)$, $\frac{1}{2}-\epsilon'$ for $do(X_1=0)$ and $\frac{1}{2}$ for all other actions. We set $\boldsymbol{q} = 0$ for $q_i \leq m$ and $\frac{1}{2}$ otherwise. Note that changing $m$ and thus $\boldsymbol{q}$ has no effect on the reward distribution.

%\eq{
%Y_t \sim \begin{cases}
%\bernoulli(\frac{1}{2}+\epsilon) & \text{if } X_1 = 1 \\
%\bernoulli(\frac{1}{2}-\frac{q_1}{1-q_1}\epsilon) & \text{otherwise}\,.
%\end{cases}
%}

@@

\begin{figure}
  \centering
  \begin{minipage}[b]{0.3\textwidth}
    \includegraphics[width=\textwidth]{exp_regret_vs_m_N50_T400_s10000_20160205_2230}
    \caption{Simple regret vs $m(\boldsymbol{q})$ for fixed horizon $T=400$ and number of variables $N = 50$}
    \label{fig:simple_vs_m}
  \end{minipage}
  \hfill
  \begin{minipage}[b]{0.3\textwidth}
    \includegraphics[width=\textwidth]{exp_regret_vs_T_N50_a4_s10000_20160204_1831}
    \caption{Simple regret vs horizon, $T$, with $N = 50$, $m=2$ and $\epsilon = \sqrt{\frac{N}{8T}}$}
    \label{fig:simple_vs_T_vary_epsilon}
  \end{minipage}
  \hfill
  \begin{minipage}[b]{0.3\textwidth}
    \includegraphics[width=\textwidth]{exp_regret_vs_T_N50_s10000_20160206_0533}
    \caption{Simple regret vs horizon, $T$, with $N = 50$, $m=2$ and fixed $\epsilon = .3$}
    \label{fig:simple_vs_T}
  \end{minipage}
\end{figure}

@@

### 2016-05-14 causal-bandit/nips_2016/general.tex
minor tweaks
first draft

\newcommand{\calP}{\mathcal P}
\newcommand{\x}{\boldsymbol{x}}
\newcommand{\Ps}{\operatorname{P}}

@@

We could naively generalize our approach for parallel bandits by observing for $T/2$ rounds, applying the truncated product factorization to write an expression for each $\P{Y|a}$ in terms of observational quantities and explicitly playing the actions for which the observational estimates were poor. However, it is no longer optimal to ignore the information we can learn about the reward for intervening on one variable from rounds in which we act on a different variable. Consider the graph in Figure \ref{fig:causalchain} and suppose each variable deterministically takes the value of its parent, $X_k = X_{k-1}$ for $k\in {2,\ldots,N}$ and $\P{X_1} = 0$. We can learn the reward for all the interventions $do(X_i = 1)$ simultaneously by selecting $do(X_1 = 1)$, but not from $do()$.

To solve the general problem we need an estimator for each action that incorporates information obtained from every other action and a way to optimally allocate samples to actions. NEED TO EXPLAIN WHY THIS PROBLEM IS HARD.

To address this difficult problem, we assume the conditional interventional distributions $\P{\parents{Y}|a}$ (but not $\P{Y|a}$) are known. These could be estimated from experimental data on the same covariates but where the outcome of interest differed, such that $Y$ was not included, or similarly from observational data subject to identifiability constraints. Let $\eta$ be a distribution on available interventions $a \in \calA$ so $\eta_a \geq 0$ and $\sum_{a \in \calA} \eta_a = 1$. Define $Q = \sum_{a \in \calA} \eta_a \P{\parents{Y}|a}$ to be the mixture distribution over the interventions with respect to $\eta$.

@@

%Firstly, the variance of the observational estimator for $a = do(X_i = j)$ can be high even if $\P{X_i = j}$ is large. For example, given the causal graph in Figure \ref{fig:causalStructure_confounded}, $\P{Y|do(X_2= j)} = \sum_{X_1}\P{X_1}\P{Y|X_1, X_2 = j}$. Suppose $X_2 = X_1$ deterministically, no matter how large $\P{X_2 = 1}$ is we will never observe $(X_2=1,X_1 = 0)$ and so cannot get a good estimate for $\P{Y|do(X_2=1)}$. Secondly it is no longer optimal

\begin{wrapfigure}[14]{r}{0.5\textwidth}
\vspace{-12pt}
\begin{minipage}{.5\textwidth}
\begin{algorithm}[H]
\caption{General Algorithm}\label{alg:general}
\begin{algorithmic}
\STATE {\bf Input:} $T$, $\eta \in [0,1]^{\calA}$, $B \in [0,\infty)^{\calA}$
\FOR{$t \in \set{1,\ldots,T}$}
\STATE Sample action $a_t = a$ with probability $\eta$
\STATE Do action $a_t$ and observe $X_t$ and $Y_t$
\ENDFOR
\FOR{$a \in \calA$}
\STATE
\eq {
\hat \mu_a =  \frac{1}{T} \sum_{t=1}^T Y_t R_a(X_t)  \ind{R_a(X_t) \leq B_a}
}
\ENDFOR
\STATE {\bf return} $\hat a^*_T = \argmax_a \hat \mu_a$
\end{algorithmic}
\end{algorithm}
\end{minipage}
\end{wrapfigure}

@@

\eq {
\hat{\mu_a} = \frac{1}{T} \sum_{t=1}^T R_a(X) Y \ind{R_a(X) \leq B_a}\,, 
}

@@

The distribution over actions, $\eta$ plays the role of allocating samples to actions and is optimized to minimize the worst-case simple regret. Abusing notation we define $m(\eta)$ by
\eq{
m(\eta) = \max_{a \in \calA} \EE_{\Ps_a}\left[\frac{\Pn{a}{\parents{Y}(X)}}{\Q{\parents{Y}(X)}}\right]\,.
}

@@

\paragraph{Choosing the Sampling Distribution} Algorithm \ref{alg:general} depends on a choice of sampling distribution $\operatorname{Q}$ that is determined by $\eta$. In light of Theorem \ref{thm:general}
a natural choice of $\eta$ is the minimiser of $m(\eta)$.
\eq{
\eta^* 
= \argmin_\eta m(\eta) = \argmin_\eta \underbrace{\max_{a \in \calA} \EE_{\Ps_a} \left[\frac{\Pn{a}{\parents{Y}(X)}}{\sum_{b \in \calA} \eta_b \Pn{b}{\parents{Y}(X)}}\right]}_{m(\eta)}\,.
}
Since the mixture of convex functions is convex and the maximum of a set of convex functions is convex, we see that $m(\eta)$ is convex (in $\eta$).
Therefore the minimisation problem may be tackled using standard techniques from convex optimisation. An interpretation of $m(\eta^*)$ is the minimum achievable worst-case variance of the importance weighted estimator. In the experimental section we present some special cases, but for now we give two simple results. The first shows that $|\calA|$ serves as an upper bound on $m(\eta^*)$.

\begin{proposition}\label{pro:m-bound}
$m(\eta^*) \leq |\calA|$.
\end{proposition}

\begin{proof}
By definition, $m(\eta^*) \leq m(\eta)$ for all $\eta$. Let $\eta_a = 1/|\calA|$. Then
\eq{
m(\eta) 
= \max_a \EE_{\Ps_a}\left[\frac{\Pn{a}{\parents{Y}(X)}}{\Q{\parents{Y}(X)}}\right] 
\leq \max_a \EE_{\Ps_a}\left[\frac{\Pn{a}{\parents{Y}(X)}}{\eta_a \Pn{a}{\parents{Y}(X)}}\right] 
\leq \max_a \EE_{\Ps_a}\left[\frac{1}{\eta_a}\right] = |\calA| \qedhere
}
\end{proof}

The second observation is that in the parallel bandit setting the $m(\eta)$ given in this section approximately coincides with the $m(\vec{q})$ in Eq.\ \ref{eq:m-simple}.
Recall in that setting that 
\eq{
\actions = \set{do()} \cup \set{do(X_i = j) : 1 \leq i \leq N \text{ and } j \in \set{0,1}}\,.
}
For $a = do()$ choose $\eta_a = 1/2$. 
Let $m = m(\vec{q})$ and note that, by definition, there are at most $m$ pairs $(i,j)$ such that $\P{X_i = j} \leq 1/m$.
Thus, for $a = do(X_i = j)$ letting $\eta_a \propto \ind{\P{X_i = j} \leq 1/m} / (2m)$ guarantees $\eta_a \ge \frac{1}{2m}$ when $\eta_a \ne 0$.
It is then easy to check that $m(\eta) \leq 2m$ using an argument like that for Proposition~\ref{pro:m-bound}.

>>>>

@@

We could naively generalize our approach for parallel bandits by observing for $T/2$ rounds, applying the truncated product factorization to write an expression for each $\P{Y|a}$ in terms of observational quantities and explicitly playing the actions for which the observational estimates were poor. However, it is no longer optimal to ignore the information we can learn about the reward for intervening on one variable from rounds in which we act on a different variable. Consider the graph in Figure \ref{fig:causalchain} and suppose each variable deterministically takes the value of its parent, $X_k = X_{k-1}$ for $k\in {2,\ldots,N}$ and $\P{X_1} = 0$. We can learn the reward for all the interventions $do(X_i = 1)$ simultaneously by selecting $do(X_1 = 1)$, but not from $do()$. In addition, variance of the observational estimator for $a = do(X_i = j)$ can be high even if $\P{X_i = j}$ is large. Given the causal graph in Figure \ref{fig:causalStructure_confounded}, $\P{Y|do(X_2= j)} = \sum_{X_1}\P{X_1}\P{Y|X_1, X_2 = j}$. Suppose $X_2 = X_1$ deterministically, no matter how large $\P{X_2 = 1}$ is we will never observe $(X_2=1,X_1 = 0)$ and so cannot get a good estimate for $\P{Y|do(X_2=1)}$.

To solve the general problem we need an estimator for each action that incorporates information obtained from every other action and a way to optimally allocate samples to actions. To address this difficult problem, we assume the conditional interventional distributions $\P{\parents{Y}|a}$ (but not $\P{Y|a}$) are known. These could be estimated from experimental data on the same covariates but where the outcome of interest differed, such that $Y$ was not included, or similarly from observational data subject to identifiability constraints. Let $\eta$ be a distribution on available interventions $a \in \calA$ so $\eta_a \geq 0$ and $\sum_{a \in \calA} \eta_a = 1$. Define $Q = \sum_{a \in \calA} \eta_a \P{\parents{Y}|a}$ to be the mixture distribution over the interventions with respect to $\eta$.

@@

\begin{wrapfigure}[12]{r}{0.5\textwidth}
\vspace{-25pt}
\begin{minipage}{.5\textwidth}
\begin{algorithm}[H]
\caption{General Algorithm}\label{alg:general}
\begin{algorithmic}
\STATE {\bf Input:} $T$, $\eta \in [0,1]^{\calA}$, $B \in [0,\infty)^{\calA}$
\FOR{$t \in \set{1,\ldots,T}$}
\STATE Sample action $a_t = a$ with probability $\eta$
\STATE Do action $a_t$ and observe $X_t$ and $Y_t$
\ENDFOR
\FOR{$a \in \calA$}
\STATE
\eq {
\hat \mu_a =  \frac{1}{T} \sum_{t=1}^T Y_t R_a(X_t)  \ind{R_a(X_t) \leq B_a}
}
\ENDFOR
\STATE {\bf return} $\hat a^*_T = \argmax_a \hat \mu_a$
\end{algorithmic}
\end{algorithm}
\end{minipage}
\end{wrapfigure}

@@

\eq {
\hat \mu_a =  \frac{1}{T} \sum_{t=1}^T Y_t R_a(X_t)  \ind{R_a(X_t) \leq B_a}\,, 
}

@@

The distribution over actions, $\eta$ plays the role of allocating samples to actions and is optimized to minimize the worst-case simple regret. Abusing notation we define $m(\eta)$ by
\eq{
m(\eta) = \max_{a \in \calA} \EEa\left[\frac{\Pn{a}{\parents{Y}(X)}}{\Q{\parents{Y}(X)}}\right]\,,\text{ where } \EEa \text{ is the expectation with respect to } \Pn{a}.
}

@@

\paragraph{Choosing the Sampling Distribution} Algorithm \ref{alg:general} depends on a choice of sampling distribution $\operatorname{Q}$ that is determined by $\eta$. In light of Theorem \ref{thm:general}
a natural choice of $\eta$ is the minimiser of $m(\eta)$.
\eq{
\eta^* 
= \argmin_\eta m(\eta) = \argmin_\eta \underbrace{\max_{a \in \calA} \EEa \left[\frac{\Pn{a}{\parents{Y}(X)}}{\sum_{b \in \calA} \eta_b \Pn{b}{\parents{Y}(X)}}\right]}_{m(\eta)}\,.
}
Since the mixture of convex functions is convex and the maximum of a set of convex functions is convex, we see that $m(\eta)$ is convex (in $\eta$).
Therefore the minimisation problem may be tackled using standard techniques from convex optimisation. An interpretation of $m(\eta^*)$ is the minimum achievable worst-case variance of the importance weighted estimator. In the experimental section we present some special cases, but for now we give two simple results. The first shows that $|\calA|$ serves as an upper bound on $m(\eta^*)$.

\begin{proposition}\label{pro:m-bound}
$m(\eta^*) \leq |\calA|$. \textit{Proof.} 
\textup{By definition, $m(\eta^*) \leq m(\eta)$ for all $\eta$. Let $\eta_a = 1/|\calA|\,\forall a$.}
\eq{
m(\eta) 
= \max_a \EEa\left[\frac{\Pn{a}{\parents{Y}(X)}}{\Q{\parents{Y}(X)}}\right] 
\leq \max_a \EEa\left[\frac{\Pn{a}{\parents{Y}(X)}}{\eta_a \Pn{a}{\parents{Y}(X)}}\right] 
= \max_a \EEa\left[\frac{1}{\eta_a}\right] = |\calA| %\qedhere
}
\end{proposition}

The second observation is that, in the parallel bandit setting, $m(\eta) \leq 2m(\boldsymbol{q})$. This is easy to see by letting $\eta_a = 1/2$ for $a = do()$ and $\eta_a = \ind{\P{X_i = j} \leq 1/m(\boldsymbol{q}} / (2m(\boldsymbol{q}))$ for the actions corresponding to $do(X_i=j)$, and applying an argument like Proposition~\ref{pro:m-bound}. The proof is in section \ref{} of the supplementary materials.

### 2016-05-14 causal-bandit/nips_2016/simple-regret.tex
first draft

The causal model for this class of problems has $N$ binary variables $\{ X_1, \ldots, X_N \}$ where each $X_i \in \{0,1\}$ are independent causes of a 
reward variable $Y \in \set{0,1}$, as shown in Figure~\ref{fig:causalStructure}.
All variables are observable and the set of allowable actions are all size 0 and size 1 interventions: $\mathcal{A} = \set{do()} \cup \set{ do(X_i = j) \colon 1 \leq i \leq N \text{ and } j \in \set{0,1}}$
In the farming example from the introduction, $X_1$ might represent temperature (\eg, $X_1=0$ for low and $X_1=1$ for high). 
The interventions $do(X_1 = 0)$ and $do(X_1 = 1)$ indicate the use of shades or heat lamps to keep the temperature low or high, respectively.

\begin{figure}
    \begin{subfigure}[b]{0.34\textwidth}
	\centering    
          \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=.45cm,
  thick,main node/.style={observed}, hidden/.style={empty},background rectangle/.style={fill=olive!45}]
%every node/.style={scale=0.6}
 %nodes
\node[main node](1){$X_{1}$};
\node[main node, right=of 1](2){$X_{2}$};
\node[hidden, right=of 2](3){$...$};
\node[main node, right=of 3](4){$X_{N}$};
\node[main node, below right=of 2](5){$Y$};
 \path[every node/.style={font=\tiny}]
    (1) edge (5)
    	(2) edge (5)
    (4) edge (5);
\end{tikzpicture}
        \caption{Parallel graph}
        \label{fig:causalStructure_confounded}
    \end{subfigure}
    \begin{subfigure}[b]{0.2\textwidth}
    \centering
        \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=.45cm,
  thick,main node/.style={observed}, hidden/.style={empty},background rectangle/.style={fill=olive!45}]
\node[main node](1){$X_1$};
\node[main node, below left=of 1](2){$X_2$};
\node[main node, below right=of 1](4){$Y$};
 \path[every node/.style={font=\tiny}]
    (1) edge (2)
    (1) edge (4)
    (2) edge (4);
\end{tikzpicture}
        \caption{Confounded graph}
        \label{fig:causalStructure_confounded}
    \end{subfigure}
    \begin{subfigure}[b]{0.4\textwidth}
    \centering
         \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=.45cm,
  thick,main node/.style={observed}, hidden/.style={empty},background rectangle/.style={fill=olive!45}]
\node[main node](1){$X_{1}$};
\node[main node, right=of 1](2){$X_{2}$};
\node[hidden, right=of 2](3){$...$};
\node[main node, right=of 3](4){$X_{N}$};
\node[main node, right=of 4](5){$Y$};
 \path[every node/.style={font=\tiny}]
    (1) edge (2)
  	(2) edge (3)
    (3) edge (4)
    (4) edge (5);
\end{tikzpicture}
        \caption{Chain graph}
        \label{fig:causalchain}
    \end{subfigure}
    \caption{Causal Models}\label{fig:animals}
\end{figure}

>>>>

The causal model for this class of problems has $N$ binary variables $\{ X_1, \ldots, X_N \}$ where each $X_i \in \{0,1\}$ are independent causes of a 
reward variable $Y \in \set{0,1}$, as shown in Figure~\ref{fig:parallel}.
All variables are observable and the set of allowable actions are all size 0 and size 1 interventions: $\mathcal{A} = \set{do()} \cup \set{ do(X_i = j) \colon 1 \leq i \leq N \text{ and } j \in \set{0,1}}$
In the farming example from the introduction, $X_1$ might represent temperature (\eg, $X_1=0$ for low and $X_1=1$ for high). 
The interventions $do(X_1 = 0)$ and $do(X_1 = 1)$ indicate the use of shades or heat lamps to keep the temperature low or high, respectively.

\begin{figure}
    \begin{subfigure}[b]{0.34\textwidth}
	\centering    
          \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=.45cm,
  thick,main node/.style={observed}, hidden/.style={empty},background rectangle/.style={fill=olive!45}]
%every node/.style={scale=0.6}
 %nodes
\node[main node](1){$X_{1}$};
\node[main node, right=of 1](2){$X_{2}$};
\node[hidden, right=of 2](3){$...$};
\node[main node, right=of 3](4){$X_{N}$};
\node[main node, below right=of 2](5){$Y$};
 \path[every node/.style={font=\tiny}]
    (1) edge (5)
    	(2) edge (5)
    (4) edge (5);
\end{tikzpicture}
        \caption{Parallel graph}
        \label{fig:parallel}
    \end{subfigure}
    \begin{subfigure}[b]{0.2\textwidth}
    \centering
        \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=.45cm,
  thick,main node/.style={observed}, hidden/.style={empty},background rectangle/.style={fill=olive!45}]
\node[main node](1){$X_1$};
\node[main node, below left=of 1](2){$X_2$};
\node[main node, below right=of 1](4){$Y$};
 \path[every node/.style={font=\tiny}]
    (1) edge (2)
    (1) edge (4)
    (2) edge (4);
\end{tikzpicture}
        \caption{Confounded graph}
        \label{fig:causalStructure_confounded}
    \end{subfigure}
    \begin{subfigure}[b]{0.4\textwidth}
    \centering
         \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=.45cm,
  thick,main node/.style={observed}, hidden/.style={empty},background rectangle/.style={fill=olive!45}]
\node[main node](1){$X_{1}$};
\node[main node, right=of 1](2){$X_{2}$};
\node[hidden, right=of 2](3){$...$};
\node[main node, right=of 3](4){$X_{N}$};
\node[main node, right=of 4](5){$Y$};
 \path[every node/.style={font=\tiny}]
    (1) edge (2)
  	(2) edge (3)
    (3) edge (4)
    (4) edge (5);
\end{tikzpicture}
        \caption{Chain graph}
        \label{fig:causalchain}
    \end{subfigure}
    \caption{Causal Models}\label{fig:causalmodels}
\end{figure}
